First and foremost, I would like to thank my advisor Dr. Herbert Tanner, for his patience and valuable guidance throughout my PhD. The knowledge and systematic outlook he instilled in me was instrumental in enabling me to finish this dissertation. I am indebted to Dr. Tanner for exposing me to a multitude of inter-disciplinary projects. This enabled me to gain valuable expertise in building robotic solutions.

Next, I would like thank members of my PhD committee: Dr. Ioannis Poulakakis, 
Dr. Christopher Rasmussen, Dr. Guoquan Huang, Dr. Arthur Trembanis and Dr. Sunil Agarwal for their comments and advice that helped me channel my dissertation in the right direction. Dr. Christopher Rasmussen also mentored me during my computer science masters and introduced me to software development using \gls{ros}. I am also grateful to Dr. Jeffrey Heinz and Dr. Adam Jardine for their valuable support in collaborative projects that exposed me to concepts in computational linguistics.

Further, I would like to thank members of \gls{cshel} at University of Delaware for helping me with data that was utilized in this dissertation. It was a great pleasure learning the way of life at sea from them (even though I got sea sick the first time I was on a ship). In addition, Dr. Art Trembanis allowed me to use the scallop data collected during his research projects. He also opened to me the facilities in College of Earth and Ocean sciences to support my experiments. I would also like to mention Hunter Brown, lab manager at \gls{rdl} for providing me access to test facilities in Lewis, Delaware. In addition, Justin H. Walker graciously provided me the annotated scallop data which was critical to my experiments. I would also like to thank Carter DuVal for building with me my first unsuccessful scrappy \gls{rov}. The \gls{rov} building continued a few years before getting to the point of a working prototype. I would also like to take this moment to thank undergraduate interns Yehan Zhang, Yang Liu and Austin Crouse for putting in countless hours into building CoopROV. Special mention to Dr. Scott Gallager and Amber York of \gls{whoi} for sharing the datasets on which Matthew Dawkings and Dr. Charles Stewart of \gls{rpi}’s Computer Science Department worked on. Thanks also to graduate students of Dr. Matthew Oliver and Dr. Mark Moline, for providing 
support during experiments at Lewis.

Moreover, various funding agencies \gls{nsf}, \gls{arl} and \gls{nih} graciously supported part of my work over the years through grants \gls{nsf} RI \#0913015, \gls{nsf} CPS \#1035577, \gls{nsf} EAGER \#1548149, \gls{nih} \#R01HD087133 and \gls{arl} MAST CTA \#W911NF-08-2-0004. I could not have done without them.

In addition, I would like to thank all my lab-mates for stimulating discussions and support through out this PhD, with special mention to Dr. Jie Fu, Dr. Shridhar Shah, Luis Valbuena Reyes, Aditya Boddu, Dr. Konstantinos Karydis, Dr. Jianxin Sun and Adam Stager. Likewise, many other university faculty, staff and fellow students also supported me in various ways. I am grateful to all of them.

Finally, I owe it to my father A.S. Kannappan, mother PL Meenakshi and brother K. Priyadharshan for their relentless moral support throughout this PhD. My girlfriend Karthika Ramanathan motivated me to get through this dissertation, I dedicate this disseration to her. In the end, I would like to raise a toast to all other friends, hiking buddies and fellow chess players who made this PhD bearable through intermittent distractions.